Solving Collective Spin Hamiltonians with Decoherence

### 2018-08-28 soc_squeezing/cumulant_expansions.tex
simplified expectation values
removed obsolete section and changed spin operator ordering

The case of the one-axis twisting is simplest to consider, as we can
write the Hamiltonian in the form
\begin{align}
  H_{\t{OAT}} = \f14 \chi \sum_{j,k} \sigma_j^z \sigma_k^z
  = \f12 \chi \sum_{j<k} \sigma_j^z \sigma_k^z + \f14 N \chi,
\end{align}
which admits an exact solution previously worked out by Foss-Feig et
al.~in ref.~\cite{foss-feig2013nonequilibrium}.  Adapting exact
expectation values with respect to the initial state $\ket{X}$
satisfying $S_x\ket{X}=S\ket{X}$, in terms of $\mu,\nu\in\set{+1,-1}$
we find that
\begin{align}
  \bk{S_+}
  &= \sum_j\bk{\sigma_j^+}
  = \f{N}{2} e^{-\gamma t/2} \Phi\p{\chi,t}^{N-1}, \label{eq:S+_OAT} \\
  \bk{S_\mu S_z}
  &= \f12\sum_j\bk{\sigma_j^\mu\sigma_j^z}
  + \f12\sum_{j\ne k} \bk{\sigma_j^\mu \sigma_k^z} \\
  &= -\f{\mu}{2}\bk{S_\mu} + \f14 N \p{N-1} e^{-\gamma t/2}
  \Psi\p{\mu\chi,t} \Phi\p{\chi,t}^{N-2}, \\
  \bk{S_\mu S_\nu}
  &= \sum_j \bk{\sigma_j^\mu \sigma_j^\nu}
  + \sum_{j\ne k} \bk{\sigma_j^\mu \sigma_k^\nu} \\
  &= \delta_{\mu,-\nu} \p{\f{N}{2} + \mu \bk{S_z}}
  + \f14 N \p{N-1} e^{-\gamma t}
  \Phi\p{\sp{\mu+\nu}\chi,t}^{N-2}, \label{eq:SS+-_OAT}
\end{align}
where
\begin{align}
  \Phi\p{\zeta,t} \equiv e^{-\gamma t/2}
  \sp{\cos\p{s_\zeta t}+\f12\gamma t~\t{sinc}\p{s_\zeta t}},
  &&
  \Psi\p{\zeta,t} \equiv e^{-\gamma t/2}
  \p{is_\zeta-\gamma/2}t~ \t{sinc}\p{s_\zeta t},
\end{align}
and $s_\zeta \equiv \zeta + i2\gamma$.  The last ingredients we need
to compute spin squeezing at any time are $\bk{S_z}$ and $\bk{S_z^2}$.
As these operators commute with the OAT Hamiltonian, their evolution
is governed entirely by decoherence, such that
\begin{align}
  \f{d}{dt}\p{S_z^n}
  = \gamma \D_-\p{S_z^n}
  = \gamma \p{S + S_z} \sum_{k=0}^{n-1}
  \p{-1}^{n-k} { n \choose k } S_z^k,
\end{align}
where $S\equiv N/2$ and $\D_-\p{S_z^n}$ is worked out in Appendix
\ref{sec:decoherence}.  In particular,
\begin{align}
  \f{d}{dt} S_z = -\gamma \p{S + S_z},
  &&
  \f{d}{dt}\p{S_z^2} = \gamma \p{S + S_z} \p{1 - 2S_z}.
\end{align}
The initial conditions $\bk{S_z}_{t=0}=0$ and $\bk{S_z^2}_{t=0}=S/2$
then imply
\begin{align}
  \bk{S_z} = S\p{e^{-\gamma t}-1},
  &&
  \bk{S_z^2} = S \p{1 - \f12 e^{-\gamma t}} e^{-\gamma t} + \bk{S_z}^2.
  \label{eq:Sz_OAT}
\end{align}
The expectation values in \eqref{eq:S+_OAT}-\eqref{eq:SS+-_OAT} and
\eqref{eq:Sz_OAT} are sufficient to compute the squeezing parameter
$\xi^2$ in \eqref{eq:squeezing} at any time throughout evolution of
the initial state $\ket{X}$ under $H_{\t{OAT}}$.

\subsection{Heisenberg systems of equations}

@@

The equations of motion for $S_\mu$, $S_\mu S_z$, and $S_\mu S_\nu$
(i.e.~where $\mu,\nu\in\set{+1,-1}$) under the OAT Hamiltonian
$H_{\t{OAT}}=\chi S_z^2$ are determined by
\begin{align}
  \f{d}{dt} \p{S_+^\ell S_-^m S_z^n}
  = i\chi \sp{S_z^2, S_+^\ell S_-^m S_z^n}_-
  + \gamma \D_-\p{S_+^\ell S_-^m S_z^n},
\end{align}
where we can use identity \eqref{eq:comm_z_mu} to simplify
\begin{align}
  \sp{S_z^2, S_+^\ell S_-^m S_z^n}_-
  = \sp{S_z^2, S_+^\ell S_-^m}_- S_z^n
  = S_+^\ell S_-^m \p{\sp{\ell-m}^2 + 2\sp{\ell-m} S_z} S_z^n,
  \label{eq:S_z^2_comm}
\end{align}
and the effect of decoherence through $\D_-\p{S_+^\ell S_-^m S_z^n}$ is
worked out in Appendix \ref{sec:decoherence}.  In order to compute
squeezing of the OAT Hamiltonian in the presence of single-spin decay,
we therefore need to solve the systems of equations defined by
\begin{align}
  \f{d}{dt} \p{S_\mu S_z^n}
  = i\chi S_\mu \p{1 + 2\mu S_z} S_z^n
  + \gamma S_\mu \sp{\p{S - \delta_{\mu,-1} + S_z} \sum_{k=0}^{n-1}
    \p{-1}^{n-k} { n \choose k } S_z^k - \f12 S_z^n}
  \label{eq:diff_mu_OAT}
\end{align}
for $\mu\in\set{+1,-1}$, and
\begin{align}
  \f{d}{dt} \p{S_+^2 S_z^n}
  = 4i\chi S_\mu S_\nu \p{1 + S_z} S_z^n
  + \gamma S_\mu S_\nu \sp{\p{S + S_z} \sum_{k=0}^{n-1}
    \p{-1}^{n-k} { n \choose k } S_z^k - S_z^n},
  \label{eq:diff_pp_OAT}
\end{align}
while $d/dt\p{S_\pm S_\mp}=0$, so
\begin{align}
  \bk{S_\pm S_\mp} = \bk{S_\pm S_\mp}_{t=0} e^{-\gamma t}.
\end{align}
The initial conditions for \eqref{eq:diff_mu_OAT} and
\eqref{eq:diff_pp_OAT} are provided in Appendix
\ref{sec:initial_conditions}.

\section{The general case}

All collective spin Hamiltonians can be written in the form
\begin{align}
  H = \sum_{p,q,r} h_{pqr} S_+^p S_-^q S_z^r.
  \label{eq:H_general}
\end{align}
The Hamiltonian for one-axis twisting with a transverse-field, for
example, is
\begin{align}
  H_{\t{TVF}}
  = \chi S_z^2 + \Omega S_x
  = \chi S_z^2 + \f{\Omega}{2} S_+ + \f{\Omega}{2} S_-,
\end{align}
while the two-axis twisting Hamiltonians about $\p{z,y}$ and $\p{y,x}$
are
\begin{align}
  H_{\t{TAT}}^{z,y}
  = \chi \p{S_z^2 - S_y^2}
  = \chi \p{S_z^2 + \f14 S_+^2 + \f14 S_-^2 - \f12 S_+ S_- + \f12 S_z},
\end{align}
and
\begin{align}
  H_{\t{TAT}}^{y,x}
  = \chi\p{S_y^2 - S_x^2}
  = -\f{\chi}{2} \p{S_+^2 + S_-^2}.
\end{align}
The evolution of a general operator $S_+^\ell S_-^m S_z^n$ under a
Hamiltonian of the form in \eqref{eq:H_general} is given by
\begin{align}
  \f{d}{dt} \p{S_+^\ell S_-^m S_z^n}
  = i\sum_{p,q,r} h_{pqr} \sp{S_+^p S_-^q S_z^r, S_+^\ell S_-^m S_z^n}_-
  + \gamma \D_-\p{S_+^\ell S_-^m S_z^n}
  = \sum_{p,q,r} T^{\ell m n}_{pqr} S_+^p S_-^q S_z^r,
\end{align}
where the commutator is given in Appendix \ref{sec:comm_general}, and
the decoherence operator in Appendix \ref{sec:decoherence}.  We can
then expand
\begin{align}
  \bk{S_+^\ell S_-^m S_z^n}
  = \sum_{k\ge0} \f{t^k}{k!}
  \bk{\f{d^k}{dt^k}\p{S_+^\ell S_-^m S_z^n}}_{t=0}
  \approx \sum_{k=0}^M \f{t^k}{k!}
  \bk{\f{d^k}{dt^k}\p{S_+^\ell S_-^m S_z^n}}_{t=0},
\end{align}
where $M$ is chosen such that
\begin{align}
  \f{t^M}{M!} \bk{\f{d^M}{dt^M}\p{S_+^\ell S_-^m S_z^n}}_{t=0}
  \ll S^{\ell+m+n}.
\end{align}
Initial conditions, i.e.~expectation values at $t=0$, are provided in
Appendix \ref{sec:initial_conditions}.
% comment on explosion of operators

@@

This text makes ubiquitous use of various spin operator identities; we
collect and derive particularly useful identities here for reference.
The elementary commutation relations between single-weight spin
operators are, with $\bmu\equiv-\mu\in\set{+1,-1}$ for brevity,
\begin{align}
  \f12 \sp{\sigma_j^z,\sigma_k^\mu}_-
  &= \delta_{jk} \mu \sigma_j^\mu,
  &
  \sp{S_z,\sigma_j^\mu}_-
  &= \f12 \sp{\sigma_j^z,S_\mu}_- = \mu \sigma_j^\mu,
  &
  \sp{S_z,S_\mu}_-
  &= \mu S_\mu,
  \label{eq:comm_z_mu_base} \\
  \sp{\sigma_j^\mu,\sigma_k^\bmu}_-
  &= \delta_{jk} \mu \sigma_j^z,
  &
  \sp{S_\mu,\sigma_j^\bmu}_-
  &= \sp{\sigma_j^\mu,S_\bmu}_- = \mu \sigma_j^z,
  &
  \sp{S_\mu,S_\bmu}_-
  &= 2 \mu S_z.
  \label{eq:comm_mu_base}
\end{align}
These relations can be used to inductively determine commutation
relations with higher-weight spin operators.  For example, we can find
that
\begin{align}
  S_z S_\mu^m
  = \p{\mu S_\mu + S_\mu S_z} S_\mu^{m-1}
  = \mu S_\mu^m + S_\mu S_z S_\mu^{m-1}
  = 2 \mu S_\mu^m + S_\mu S_z S_\mu^{m-2}
  = S_\mu^m \p{\mu m + S_z},
  \label{eq:push_z_mu_single}
\end{align}
and in turn
\begin{align}
  S_z S_\mu^m S_\bmu^n
  = S_\mu^m \p{\mu m + S_z} S_\bmu^n
  = S_\mu^m S_\bmu^n \p{\mu\sp{m - n} + S_z}.
\end{align}
Repeating this process $\ell$ times tells us
\begin{align}
  S_z^\ell S_\mu^m S_\bmu^n
  = S_\mu^m S_\bmu^n \p{\mu\sp{m - n} + S_z}^\ell,
  \label{eq:push_z_mu}
\end{align}
which implies
\begin{align}
  \sp{S_z^\ell, S_\mu^m S_\bmu^n}_-
  = S_\mu^m S_\bmu^n \sp{\mu^\ell \p{\mu\sp{m - n} + S_z}^\ell - S_z^\ell}
  = S_\mu^m S_\bmu^n  \sum_{k=0}^{\ell-1} {\ell \choose k}
  \mu^{\ell-k} \p{m-n}^{\ell-k} S_z^k.
  \label{eq:comm_z_mu}
\end{align}
Similarly finding commutation relations between high-weight transverse
spin operators, i.e.~$S_\mu^m$ and $S_\bmu^n$, turns out to be
considerably more difficult than the cases we have worked out thus
far.  We therefore save this work for Appendix
\ref{sec:comm_transverse}.

@@

(arbitrarily) choose to impose an ordering with all $S_z$ operators on
the right.  This choice of ordering compels us to expand

@@

Observing that we always have
$0\le k_1\le k_2\le\cdots\le k_p\le m-p-1$ in \eqref{eq:E_mn_p}, we
can rearrange the order of the sums and relabel $k_p\to\ell$ to get
\begin{align}
  E_{mn;\mu}^{(p)}
  = \f{n!}{\p{n-p}!}
  \sum_{\ell=0}^{m-p-1} D_{m-p,n-p;\mu}^{(\ell)} \p{2\ell+F_{np;\mu}}
  \sum_{\p{\v k,p,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}},
  \label{eq:E_mn_p_sum}
\end{align}
where for shorthand we define
\begin{align}
  F_{np;\mu} \equiv p - n + 2\mu S_z,
  &&
  \sum_{\p{\v k,p,\ell}} X \equiv
  \sum_{k_1=0}^\ell \sum_{k_2=k_1}^\ell
  \cdots \sum_{k_{p-1}=k_{p-2}}^\ell X.
\end{align}
We now define
\begin{align}
  f_{np\ell;\mu}\p{k,q} \equiv \p{\ell-k+q} \p{\ell+k-q+F_{np;\mu}},
\end{align}
and evaluate sums successively over $k_{p-1},k_{p-2},\cdots,k_1$,
finding
\begin{align}
  \sum_{\p{\v k,p,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
  &= \sum_{\p{\v k,p-1,\ell}}
  \prod_{j=2}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
  f_{np\ell;\mu}\p{k_{p-2},1} \\
  &= \f1{r!} \sum_{\p{\v k,p-r,\ell}}
  \prod_{j=1+r}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
  \prod_{q=1}^r f_{np\ell;\mu}\p{k_{p-1-r},q} \\
  &= \f1{\p{p-1}!} \prod_{q=1}^{p-1} f_{np\ell;\mu}\p{0,q} \\
  &= { \ell + p - 1 \choose p - 1 }
  \prod_{q=1}^{p-1} \p{\ell-q+F_{np;\mu}}.
\end{align}
Substitution of this result together with $D_{m-p,n-p;\mu}^{(\ell)}$
with \eqref{eq:D_mn_k} into \eqref{eq:E_mn_p_sum} then gives us
\begin{align}
  E_{mn;\mu}^{(p)}
  = \f{n!}{\p{n-p-1}!} S_\bmu^{n-p-1} S_\mu^{m-p-1} G_{mnp;\mu}
\end{align}
with
\begin{align}
  G_{mnp;\mu}
  &\equiv \sum_{\ell=0}^{m-p-1} { \ell + p - 1 \choose p - 1 }
  \p{m-p-\ell} \p{\ell+m-p+F_{np;\mu}}
  \p{2\ell + F_{np;\mu}}
  \prod_{q=1}^{p-1} \p{\ell-q+F_{np;\mu}} \\
  &= { m \choose p + 1 } \prod_{q=0}^p \p{m-p-q+F_{np;\mu}}.
\end{align}
We can further expand
\begin{align}
  \prod_{q=0}^p \p{m-p-q+F_{np;\mu}}
  = \prod_{q=0}^p \p{m-n-q+2\mu S_z}
  = \sum_{q=0}^{p+1} \p{-1}^{p+1-q} { p+1 \brack q } \p{m-n+2\mu S_z}^q,
\end{align}
where ${ n \brack k }$ is an unsigned Stirling number of the first
kind, and finally
\begin{align}
  \sum_{q=0}^p \p{-1}^{p-q} { p \brack q } \p{m-n+2\mu S_z}^q
  &= \sum_{q=0}^p \p{-1}^{p-q} { p \brack q } \sum_{\ell=0}^q
  { q \choose \ell } \p{m-n}^{q-\ell} \p{2\mu S_z}^\ell \\
  &= \sum_{\ell=0}^p \p{2\mu}^\ell \sum_{q=\ell}^p \p{-1}^{p-q}
  { p \brack q } { q \choose \ell } \p{m-n}^{q-\ell} S_z^\ell.
\end{align}
Putting everything together, we finally have
\begin{align}
  E_{mn;\mu}^{(p-1)}
  = p! { m \choose p } { n \choose p }
  S_\bmu^{n-p} S_\mu^{m-p} \sum_{\ell=0}^p
  \mu^\ell \epsilon_{mnp}^{(\ell)} S_z^\ell,
\end{align}
with
\begin{align}
  \epsilon_{mnp}^{(\ell)}
  \equiv 2^\ell \sum_{q=\ell}^p \p{-1}^{p-q}
  { p \brack q } { q \choose \ell } \p{m-n}^{q-\ell},
  \label{eq:epsilon}
\end{align}
where in this final form $E_{mn;\mu}^{(0)} = D_{mn;\mu}$, which
together with \eqref{eq:C_mn_E} implies that
\begin{align}
  \sp{S_\mu^m, S_\bmu^n}_-
  = \sum_{p=1}^{\min\set{m,n}}
  p! { m \choose p } { n \choose p } S_\bmu^{n-p} S_\mu^{m-p}
  \sum_{\ell=0}^p \mu^\ell \epsilon_{mnp}^{(\ell)} S_z^\ell,
  \label{eq:comm_mu}
\end{align}
and
\begin{align}
  S_\mu^m S_\bmu^n
  = \sum_{p=0}^{\min\set{m,n}}
  p! { m \choose p } { n \choose p } S_\bmu^{n-p} S_\mu^{m-p}
  \sum_{\ell=0}^p \mu^\ell \epsilon_{mnp}^{(\ell)} S_z^\ell.
  \label{eq:push_mu_nu}
\end{align}

@@

The decoherence operator for uncorrelated decay ($\mu=-1$) or
excitation ($\mu=1$) of individual spins is, with $\bmu\equiv-\mu$,
\begin{align}
  \D_\mu\p{\O}
  \equiv \sum_j\p{\sigma_j^\bmu \O \sigma_j^\mu
    - \f12\sp{\sigma_j^\bmu \sigma_j^\mu,\O}_+},
\end{align}
where
\begin{align}
  \sum_j \f12\sp{\sigma_j^\bmu \sigma_j^\mu,\O}_+
  = \f12\sp{S-\mu S_z,\O}_+
  = S\O - \f{\mu}{2}\sp{S_z,\O}_+
  = \O\p{S-\mu S_z} - \f{\mu}{2}\sp{S_z,\O}_-,
\end{align}
so
\begin{align}
  \D_\mu\p{\O}
  = \sum_j \sigma_j^\bmu \O \sigma_j^\mu
  - \O\p{S - \mu S_z} + \f{\mu}{2} \sp{S_z,\O}_-.
\end{align}

>>>>

The case of the one-axis twisting is simplest to consider, as we can
write the Hamiltonian in the form
\begin{align}
  H_{\t{OAT}} = \f14 \chi \sum_{j,k} \sigma_j^z \sigma_k^z
  = \f12 \chi \sum_{j<k} \sigma_j^z \sigma_k^z + \f14 N \chi,
\end{align}
which admits an exact solution previously worked out by Foss-Feig et
al.~in ref.~\cite{foss-feig2013nonequilibrium}.  Adapting exact
expectation values with respect to the initial state $\ket{X}$
satisfying $S_x\ket{X}=S\ket{X}$, in terms of $\mu,\nu\in\set{+1,-1}$
we find that
\begin{align}
  \bk{S_+}
  &= \sum_j\bk{\sigma_j^+}
  = \f{N}{2} e^{-\gamma t/2} \Phi\p{\chi,t}^{N-1}, \label{eq:S+_OAT} \\
  \bk{S_\mu S_z}
  &= \f12\sum_j\bk{\sigma_j^\mu\sigma_j^z}
  + \f12\sum_{j\ne k} \bk{\sigma_j^\mu \sigma_k^z} \\
  &= -\f{\mu}{2}\bk{S_\mu} + \f14 N \p{N-1} e^{-\gamma t/2}
  \Psi\p{\mu\chi,t} \Phi\p{\chi,t}^{N-2}, \\
  \bk{S_\mu S_\nu}
  &= \sum_j \bk{\sigma_j^\mu \sigma_j^\nu}
  + \sum_{j\ne k} \bk{\sigma_j^\mu \sigma_k^\nu} \\
  &= \delta_{\mu,-\nu} \p{\f{N}{2} + \mu \bk{S_z}}
  + \f14 N \p{N-1} e^{-\gamma t}
  \Phi\p{\sp{\mu+\nu}\chi,t}^{N-2}, \label{eq:SS+-_OAT}
\end{align}
where
\begin{align}
  \Phi\p{\xi,t} \equiv e^{-\gamma t/2}
  \sp{\cos\p{s_\xi t}+\f12\gamma t~\t{sinc}\p{s_\xi t}},
  &&
  \Psi\p{\xi,t} \equiv e^{-\gamma t/2}
  \p{is_\xi-\gamma/2}t~ \t{sinc}\p{s_\xi t},
\end{align}
and $s_\xi \equiv \xi + i2\gamma$.  The last ingredients we need
to compute spin squeezing at any time are $\bk{S_z}$ and $\bk{S_z^2}$.
As these operators commute with the OAT Hamiltonian, their evolution
is governed entirely by decoherence, such that
\begin{align}
  \f{d}{dt}\p{S_z^n}
  = \gamma \D_-\p{S_z^n}
  = \gamma \p{S + S_z} \sum_{k=0}^{n-1}
  \p{-1}^{n-k} { n \choose k } S_z^k,
\end{align}
where $S\equiv N/2$ and $\D_-\p{S_z^n}$ is worked out in Appendix
\ref{sec:decoherence}.  In particular,
\begin{align}
  \f{d}{dt} S_z = -\gamma \p{S + S_z},
  &&
  \f{d}{dt}\p{S_z^2} = \gamma \p{S + S_z} \p{1 - 2S_z}.
\end{align}
The initial conditions $\bk{S_z}_{t=0}=0$ and $\bk{S_z^2}_{t=0}=S/2$
then imply
\begin{align}
  \bk{S_z} = S\p{e^{-\gamma t}-1},
  &&
  \bk{S_z^2} = S \p{1 - \f12 e^{-\gamma t}} e^{-\gamma t} + \bk{S_z}^2.
  \label{eq:Sz_OAT}
\end{align}
The expectation values in \eqref{eq:S+_OAT}-\eqref{eq:SS+-_OAT} and
\eqref{eq:Sz_OAT} are sufficient to compute the squeezing parameter
$\xi^2$ in \eqref{eq:squeezing} at any time throughout evolution of
the initial state $\ket{X}:S_x\ket{X}=S\ket{X}$ under $H_{\t{OAT}}$.

\section{Cumulant expansion and Heisenberg systems of equations}

@@

All collective spin Hamiltonians can be written in the form
\begin{align}
  H = \sum_{p,q,r} h_{pqr} S_+^p S_z^q S_-^r.
  \label{eq:H_general}
\end{align}
The Hamiltonians for one-axis twisting with a transverse-field, for
example, is
\begin{align}
  H_{\t{TVF}}
  = \chi S_z^2 + \Omega S_x
  = \chi S_z^2 + \f{\Omega}{2} S_+ + \f{\Omega}{2} S_-,
\end{align}
while the two-axis twisting Hamiltonians about $\p{z,y}$ and $\p{y,x}$
are
\begin{align}
  H_{\t{TAT}}^{z,y}
  = \chi \p{S_z^2 - S_y^2}
  = \chi \p{S_z^2 + \f14 S_+^2 + \f14 S_-^2 - \f12 S_+ S_- + \f12 S_z},
\end{align}
and
\begin{align}
  H_{\t{TAT}}^{y,x}
  = \chi\p{S_y^2 - S_x^2}
  = -\f{\chi}{2} \p{S_+^2 + S_-^2}.
\end{align}
The evolution of a general operator $S_+^\ell S_-^m S_z^n$ under a
Hamiltonian of the form in \eqref{eq:H_general} is given by
\begin{align}
  \f{d}{dt} \p{S_+^\ell S_z^m S_-^n}
  = i\sum_{p,q,r} h_{pqr} \sp{S_+^p S_z^q S_-^r, S_+^\ell S_z^m S_-^n}_-
  + \gamma \D_-\p{S_+^\ell S_z^m S_-^n}
  \equiv \sum_{p,q,r} T^{\ell m n}_{pqr} S_+^p S_z^q S_-^r,
\end{align}
where the commutator is worked out in Appendix \ref{sec:comm_general},
and the decoherence operator in Appendix \ref{sec:decoherence}.  We
can then expand
\begin{align}
  \bk{S_+^\ell S_z^m S_-^n}
  = \sum_{k\ge0} \f{t^k}{k!}
  \bk{\f{d^k}{dt^k}\p{S_+^\ell S_z^m S_-^n}}_{t=0}
  \approx \sum_{k=0}^{M-1} \f{t^k}{k!}
  \bk{\f{d^k}{dt^k}\p{S_+^\ell S_z^m S_-^n}}_{t=0},
\end{align}
where $M$ is chosen such that
\begin{align}
  \f{t^M}{M!} \bk{\f{d^M}{dt^M}\p{S_+^\ell S_z^m S_-^n}}_{t=0}
  \ll S^{\ell+m+n}.
\end{align}
Initial conditions, i.e.~expectation values at $t=0$, are provided in
Appendix \ref{sec:initial_conditions}.
% comment on explosion of operators

@@

This text makes ubiquitous use of various spin operator identities; we
collect and derive particularly useful identities here for reference.
The elementary commutation relations between single-weight spin
operators are, with $\bmu\equiv-\mu\in\set{+1,-1}$ for brevity,
\begin{align}
  \f12 \sp{\sigma_j^z,\sigma_k^\mu}_-
  &= \delta_{jk} \mu \sigma_j^\mu,
  &
  \sp{S_z,\sigma_j^\mu}_-
  &= \f12 \sp{\sigma_j^z,S_\mu}_- = \mu \sigma_j^\mu,
  &
  \sp{S_z,S_\mu}_-
  &= \mu S_\mu,
  \label{eq:comm_z_mu_base} \\
  \sp{\sigma_j^\mu,\sigma_k^\bmu}_-
  &= \delta_{jk} \mu \sigma_j^z,
  &
  \sp{S_\mu,\sigma_j^\bmu}_-
  &= \sp{\sigma_j^\mu,S_\bmu}_- = \mu \sigma_j^z,
  &
  \sp{S_\mu,S_\bmu}_-
  &= 2 \mu S_z.
  \label{eq:comm_mu_base}
\end{align}
These relations can be used to inductively determine commutation
relations with higher-weight spin operators.  For example, we can find
that
\begin{align}
  S_z S_\mu^m
  = \p{\mu S_\mu + S_\mu S_z} S_\mu^{m-1}
  = \mu S_\mu^m + S_\mu S_z S_\mu^{m-1}
  = 2 \mu S_\mu^m + S_\mu S_z S_\mu^{m-2}
  = S_\mu^m \p{\mu m + S_z}.
  \label{eq:push_z_mu_single}
\end{align}
Repeating this process $\ell$ times tells us
\begin{align}
  S_z^\ell S_\mu^m
  = S_\mu^m \p{\mu m + S_z}^\ell,
  \label{eq:push_z_mu}
\end{align}
which implies
\begin{align}
  \sp{S_z^\ell, S_\mu^m}_-
  = S_\mu^m \sp{\p{\mu m + S_z}^\ell - S_z^\ell}
  = S_\mu^m \sum_{k=0}^{\ell-1} {\ell \choose k} \p{\mu m}^{\ell-k} S_z^k.
  \label{eq:comm_z_mu}
\end{align}
Similarly finding commutation relations between high-weight transverse
spin operators, i.e.~$S_\mu^m$ and $S_\bmu^n$, turns out to be
considerably more difficult than the cases we have worked out thus
far.  We therefore save this work for Appendix
\ref{sec:comm_transverse}.

@@

choose to impose an ordering for which all collective spin operators
take the form $S_\bmu^\ell S_\mu^m S_z^n$, as such an ordering will
prove convenient for the calculations in this section.  This choice of
ordering compels us to expand

@@

Observing that we always have
$0\le k_1\le k_2\le\cdots\le k_p\le m-p-1$ in \eqref{eq:E_mn_p}, we
can rearrange the order of the sums and relabel $k_p\to\ell$ to get
  \sum_{\ell=0}^{m-p-1} D_{m-p,n-p;\mu}^{(\ell)} \p{2\ell+F_{np;\mu}}
  \sum_{\p{\v k,p,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}},
  \label{eq:E_mn_p_sum}
where for shorthand we define
  F_{np;\mu} \equiv p - n + 2\mu S_z,
  \sum_{\p{\v k,p,\ell}} X \equiv
  \sum_{k_1=0}^\ell \sum_{k_2=k_1}^\ell
  \cdots \sum_{k_{p-1}=k_{p-2}}^\ell X.
We now define
  f_{np\ell;\mu}\p{k,q} \equiv \p{\ell-k+q} \p{\ell+k-q+F_{np;\mu}},
and evaluate sums successively over $k_{p-1},k_{p-2},\cdots,k_1$,
finding
  \sum_{\p{\v k,p,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
  &= \sum_{\p{\v k,p-1,\ell}}
  \prod_{j=2}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
  f_{np\ell;\mu}\p{k_{p-2},1} \\
  &= \f1{r!} \sum_{\p{\v k,p-r,\ell}}
  \prod_{j=1+r}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
  \prod_{q=1}^r f_{np\ell;\mu}\p{k_{p-1-r},q} \\
  &= \f1{\p{p-1}!} \prod_{q=1}^{p-1} f_{np\ell;\mu}\p{0,q} \\
  &= { \ell + p - 1 \choose p - 1 }
  \prod_{q=1}^{p-1} \p{\ell-q+F_{np;\mu}}.
Substitution of this result together with $D_{m-p,n-p;\mu}^{(\ell)}$
with \eqref{eq:D_mn_k} into \eqref{eq:E_mn_p_sum} then gives us
  = \f{n!}{\p{n-p-1}!} S_\bmu^{n-p-1} S_\mu^{m-p-1} G_{mnp;\mu}
with
  G_{mnp;\mu}
  &\equiv \sum_{\ell=0}^{m-p-1} { \ell + p - 1 \choose p - 1 }
  \p{m-p-\ell} \p{\ell+m-p+F_{np;\mu}}
  \p{2\ell + F_{np;\mu}}
  \prod_{q=1}^{p-1} \p{\ell-q+F_{np;\mu}} \\
  &= { m \choose p + 1 } \prod_{q=0}^p \p{m-p-q+F_{np;\mu}}.
We can further expand
  \prod_{q=0}^p \p{m-p-q+F_{np;\mu}}
  = \prod_{q=0}^p \p{m-n-q+2\mu S_z}
  = \sum_{q=0}^{p+1} \p{-1}^{p+1-q} { p+1 \brack q } \p{m-n+2\mu S_z}^q,
where ${ n \brack k }$ is an unsigned Stirling number of the first
kind, and finally
  \sum_{q=0}^p \p{-1}^{p-q} { p \brack q } \p{m-n+2\mu S_z}^q
  &= \sum_{q=0}^p \p{-1}^{p-q} { p \brack q } \sum_{\ell=0}^q
  { q \choose \ell } \p{m-n}^{q-\ell} \p{2\mu S_z}^\ell \\
  &= \sum_{\ell=0}^p \p{2\mu}^\ell \sum_{q=\ell}^p \p{-1}^{p-q}
  { p \brack q } { q \choose \ell } \p{m-n}^{q-\ell} S_z^\ell.
Putting everything together, we finally have
  E_{mn;\mu}^{(p-1)}
  = p! { m \choose p } { n \choose p }
  S_\bmu^{n-p} S_\mu^{m-p} \sum_{\ell=0}^p
  \mu^\ell \epsilon_{mnp}^{(\ell)} S_z^\ell,
with
  \epsilon_{mnp}^{(\ell)}
  \equiv 2^\ell \sum_{q=\ell}^p \p{-1}^{p-q}
  { p \brack q } { q \choose \ell } \p{m-n}^{q-\ell},
  \label{eq:eps_mnp}
where in this final form $E_{mn;\mu}^{(0)} = D_{mn;\mu}$, which
together with \eqref{eq:C_mn_E} implies that
  = \sum_{p=1}^{\min\set{m,n}}
  p! { m \choose p } { n \choose p } S_\bmu^{n-p} S_\mu^{m-p}
  \sum_{\ell=0}^p \mu^\ell \epsilon_{mnp}^{(\ell)} S_z^\ell,
  \label{eq:comm_mu}
  = \sum_{p=0}^{\min\set{m,n}}
  p! { m \choose p } { n \choose p } S_\bmu^{n-p} S_\mu^{m-p}
  \sum_{\ell=0}^p \mu^\ell \epsilon_{mnp}^{(\ell)} S_z^\ell,
  \label{eq:push_mu_nu}
If we wish to order collective spin operators with $S_z$ in between
$S_\bmu$ and $S_\mu$, then we find
  = \sum_{p=0}^{\min\set{m,n}} p! { m \choose p } { n \choose p }
  S_\bmu^{n-p} Z_{mn;\mu}^{(p)} S_\mu^{m-p},
where
  Z_{mn;\mu}^{(p)}
  \equiv \sum_{\ell=0}^p \mu^\ell \epsilon_{mnp}^{(\ell)}
  \p{-\mu\sp{m-p} + S_z}^\ell
  = \sum_{q=0}^p \mu^q \xi_{mnp}^{(q)} S_z^q,
  \label{eq:Z_mn}
with
  \xi_{mnp}^{(q)}
  \equiv \sum_{\ell=q}^p \p{-1}^{\ell-q}
  \epsilon_{mnp}^{(\ell)} { \ell \choose q } \p{m-p}^{\ell-q}
  \label{eq:xi_mnp}
in terms of $\epsilon_{mnp}^{(\ell)}$ as defined in
\eqref{eq:eps_mnp}.

@@

The decoherence operator for uncorrelated decay ($\mu=-1$) or
excitation ($\mu=1$) of individual spins is, with $\bmu\equiv-\mu$,
\begin{align}
  \D_\mu\p{\O}
  \equiv \sum_j\p{\sigma_j^\bmu \O \sigma_j^\mu
    - \f12\sp{\sigma_j^\bmu \sigma_j^\mu,\O}_+},
\end{align}
where
\begin{align}
  -\f12 \sp{\sigma_j^\bmu \sigma_j^\mu,\O}_+
  = -\f12\sp{\p{S-\mu S_z}, \O}_+
  = -S\O + \f{\mu}{2} \sp{S_z, \O}_+,
\end{align}
so
\begin{align}
  \D_\mu\p{\O}
  = \sum_j\sigma_j^\bmu \O \sigma_j^\mu
    - S \O + \f{\mu}{2} \sp{S_z, \O}_+.
  \label{eq:D_mu}
\end{align}
